---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# peridefs

<!-- badges: start -->
<!-- badges: end -->

Provides code sets (ICD-9, ICD-10, and HCPCS/CPT) to define conditions such as
hypertension, diabetes, and atherosclerotic cardiovascular disease in real world
claims data. It also contains generic names for common drug types (e.g.,
anti-diabetes). The purpose of the `peridefs` package is to provide safe and
easy access to these validated definitions.

## Why peridefs?
- **Dynamic and safe**: each condition and drug class can be updated over time,
but those updates will not break existing code or force that code to use the
updated definitions.
- **Built by experts**: every code set is designed by Dr Ligong Chen and Hong
Zhao, who have decades of experience working with claims data and have developed
each code set following a standardized, validated procedure.
- **Adaptable**: codes are stored in pre-built specs for standard conditions,
and they can be modified through a built-in API that allows codes to be removed
or added. We do not recommend changing our specs, but we do want
  that to be an option in case it's needed.

## Installation

``` r
# Install from GitHub
pak::pak("perisphere-rwe/peridefs")
```

## Quick start

A quick look at how you can get relevant information from `peridefs` family of `get_` functions

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(peridefs)

# Retrieve ICD-10 hypertension codes as a tibble
get_htn_v1_codes(code_type = "dx_icd10", format = "tibble")

# Complement with icd package to look up code explanations
library(icd) # for as.icd9 / as.icd10 / explain_code functions
library(tidyverse) # for map / set_names / enframe functions

get_htn_v1_codes(code_type = 'dx_icd10')[[1]] %>% 
  as.icd10() %>% 
  set_names() %>% 
  map_chr(~explain_code(.x)) %>% 
  enframe(name = 'code', value = 'explanation')

# Retrieve generic drug names for ACE inhibitors (version 2)
get_antihypertensive_generics(component = 'acei_v2')
```

## Specs

`peridefs` is built on an R6 back-end that provides `specs` for conditions and drug classes. You can work directly with our exported specs if the `get_` functions don't provide exactly what you're looking for

```{r}
# the spec objects all have a built-in print method that
# gives a helpful summary and recommended method for 
# defining the given condition.
spec_htn_v1

# just like get_htn_v1_codes() above
spec_htn_v1$get_codes()
```

