---
title: "Creating Custom Specs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Custom Specs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

```{r load}
library(peridefs)
```

All bundled specs are immutable — they are read-only data objects that ship
with the package. When you need a code set that differs from a bundled spec
(e.g., a project-specific algorithm, a subset of codes, or a study-defined
version), `peridefs` provides functions to create and modify specs without
touching the originals.

## Creating a new CodeSpec

`code_spec()` creates a new [`CodeSpec`](../reference/CodeSpec.html) from
scratch. Provide a short `condition` ID, a `label`, and a named list of code
vectors:

```{r code-spec}
my_htn <- code_spec(
  condition = "my_htn",
  label     = "Hypertension (study version)",
  version   = "v1",
  codes     = list(
    dx_icd9  = c("4010", "4011", "4019"),
    dx_icd10 = c("I10")
  )
)

my_htn
```

The resulting spec object has `$get_codes()` and `$get_defs()` methods:

```{r code-spec-get}
my_htn$get_codes(code_type = "dx_icd10")
```

## Creating a new DrugSpec

`drug_spec()` creates a [`DrugSpec`](../reference/DrugSpec.html) from a vector
of generic names:

```{r drug-spec}
my_thiazide <- drug_spec(
  drug_class    = "my_thiazide",
  label         = "Thiazide Diuretics (study subset)",
  version       = "v1",
  generic_names = c("HYDROCHLOROTHIAZIDE", "CHLORTHALIDONE", "INDAPAMIDE")
)

my_thiazide$get_generics()
```

## Adding codes to a bundled spec

`add_codes()` returns a modified deep clone of a spec with additional codes
merged in. The `...` argument names must be valid code-set keys. The original
is never modified:

```{r add-codes}
# Inspect available keys
names(spec_htn_v1$get_codes())

# Add a code and confirm the clone grew; original is unchanged
htn_plus <- add_codes(spec_htn_v1, dx_icd10 = c("I99"))

length(htn_plus$get_codes(code_type = "dx_icd10")$dx_icd10)  # clone
length(spec_htn_v1$get_codes(code_type = "dx_icd10")$dx_icd10)  # original
```

By default, added codes are flagged for both `condition` and `outcome`.
Use `variable_type` to restrict membership:

```{r add-codes-outcome}
htn_outcome_only <- add_codes(
  spec_htn_v1,
  variable_type = "outcome",
  dx_icd10      = c("I99")
)
```

## Removing codes from a bundled spec

`remove_codes()` returns a clone with the specified codes excluded:

```{r remove-codes}
htn_trimmed <- remove_codes(spec_htn_v1, dx_icd9 = c("4019"))

# Code removed from clone; original is unchanged
"4019" %in% htn_trimmed$get_codes(code_type = "dx_icd9")$dx_icd9
"4019" %in% spec_htn_v1$get_codes(code_type = "dx_icd9")$dx_icd9
```

## Modifying a CodeSpec

`modify_code_spec()` can update the label, narrative definitions, or replace
code sets:

```{r modify-code-spec}
htn_revised <- modify_code_spec(
  spec  = spec_htn_v1,
  label = "Hypertension (revised)",
  defs  = list(
    condition = c("i" = "Study-specific hypertension definition."),
    outcome   = NULL
  )
)

htn_revised
```

To create a **new version**, construct a new `code_spec()` object directly
rather than modifying an existing one. This keeps each version explicit and
self-contained:

```{r new-version}
spec_my_htn_v2 <- code_spec(
  condition = "my_htn",
  label     = "Hypertension (study v2)",
  version   = "v2",
  defs      = list(
    condition = c("i" = "Expanded definition including I11.x and I13.x."),
    outcome   = NULL
  ),
  codes = list(
    dx_icd10 = c("I10", "I110", "I119", "I130", "I132")
  )
)

spec_my_htn_v2
```

## Modifying a DrugSpec

`modify_drug_spec()` mirrors `modify_code_spec()` for drug classes. To work
with a specific component from a composite, extract it first:

```{r modify-drug-spec}
# Extract a leaf spec from a composite
acei_v1 <- spec_antihypertensive$components()$acei_v1

# Build a study-specific variant
my_acei <- modify_drug_spec(
  spec          = acei_v1,
  label         = "ACE Inhibitors (study subset)",
  generic_names = c("LISINOPRIL", "RAMIPRIL", "ENALAPRIL")
)

my_acei$get_generics()
```

## Immutability

All custom-spec functions return **deep clones**. Bundled spec objects
(`spec_htn_v1`, `spec_antihypertensive`, etc.) are never modified in place,
so you can safely build study-specific variants without affecting other code
in your project that relies on the standard definitions.
