---
title: "Working with Drug Specs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Drug Specs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

```{r load}
library(peridefs)
```

All drug specs in `peridefs` are **composites** — they contain named component
specs (individual drug classes or versions) rather than a flat GNN list. The
`component` argument is required on every drug `get_*` function.

| Function | Returns |
|----------|---------|
| `get_<drug>_generics(component = ...)` | Character vector of generic nonproprietary names (GNNs) |
| `get_<drug>_codes(component = ...)` | Character vector of NDC codes (typically empty) |
| `get_<drug>_defs(component = ...)` | Narrative description of the component's drug class definition |

Print any composite spec to see its available component names:

```{r antihtn-spec}
spec_antihypertensive
```

## Retrieving GNNs for a specific component

Pass the versioned component name to `get_*_generics()`:

```{r acei-v1}
get_antihypertensive_generics(component = "acei_v1")
```

Use `component = "all"` to retrieve the union of every component's GNNs:

```{r antihtn-all}
get_antihypertensive_generics(component = "all") |> length()
```

### Comparing versions

Many drug classes have two versions — v1 from the Perisphere medication list
and v2 from First DataBank (FDB), which may add spelling variants:

```{r acei-compare}
# v1 — Perisphere source
get_antihypertensive_generics(component = "acei_v1")

# v2 — FDB (adds FOSINIPRIL and MOEXEPRIL variants)
get_antihypertensive_generics(component = "acei_v2")
```

## Algorithm definitions

Each component carries a narrative description of how that drug class was
identified:

```{r acei-defs}
get_antihypertensive_defs(component = "acei_v1")
```

## NDC codes

NDC code vectors are currently empty for drug classes. They can be added in a
future update.  The `get_<drug>_codes()` function is reserved for future use or
for custom specs you build with `drug_spec()`. The primary identification
mechanism at the moment is through generic names.

## Antidiabetic drug classes

`spec_antidiabetic` groups all oral antidiabetic agents and insulin:

```{r antidiab-spec}
spec_antidiabetic
```

```{r glp1}
get_antidiabetic_generics(component = "glp1_v1")
```

```{r sglt2}
get_antidiabetic_generics(component = "sglt2_v1")
```

## Creating your own drug spec

Use `drug_spec()` to define a custom drug class, or `modify_drug_spec()` to
extend an existing component. See `vignette("custom_specs")` for details.
