---
title: "Working with Drug Specs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Drug Specs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

```{r load}
library(peridefs)
```

Most drug specs in `peridefs` are **composites** — they contain named component
specs (individual drug classes or versions) rather than a flat GNN list. The
`component` argument is optional on every drug `get_*` function; omit it (or
pass `"all"`) to retrieve every component at once.

| Function | Returns |
|----------|---------|
| `get_<drug>_generics(component = ...)` | Tibble of generic (and brand, as a list-column) names, with `priority`, `class`, and `version` columns |
| `get_<drug>_defs(component = ...)` | Narrative description of the component's drug class definition |

Print any composite spec to see its available component names:

```{r antihtn-spec}
spec_antihypertensive
```

## Retrieving GNNs for a specific component

Pass the versioned component name to `get_*_generics()`:

```{r acei-v1}
get_antihypertensive_generics(component = "acei_v1")
```

Omit `component` (or pass `"all"`) to retrieve every component's GNNs at once,
distinguished by the `class` and `version` columns:

```{r antihtn-all}
get_antihypertensive_generics() |> nrow()
```

### Comparing versions

Many drug classes have two versions — v1 from the Perisphere medication list
and v2 from First DataBank (FDB), which may add spelling variants:

```{r acei-compare}
# v1 — Perisphere source
get_antihypertensive_generics(component = "acei_v1")

# v2 — FDB (adds FOSINIPRIL and MOEXEPRIL variants)
get_antihypertensive_generics(component = "acei_v2")
```

## Algorithm definitions

Each component carries a narrative description of how that drug class was
identified:

```{r acei-defs}
get_antihypertensive_defs(component = "acei_v1")
```

## Confidence tiers

The `priority` column distinguishes how confident we are that a generic name
belongs in a drug class definition: `1` (core) is the default and most
conservative tier, `2` (probable) covers generics with more than one
indication, and `3` (cautious) covers generics without their expected
indication in the US. Pass `priority = 1:3` to widen the result and include
lower-confidence matches.

## Brand names

Some generics genuinely correspond to more than one brand product — most
often because the same `GNRC_NM` covers formulations approved for different
indications (e.g., semaglutide as Ozempic for type 2 diabetes vs. Wegovy for
chronic weight management). Because a single generic can map to zero, one, or
multiple brands, `brand` is stored as a **list-column**: each row's `brand`
value is a character vector, empty (`character(0)`) when no brand is on
record.

Build a small example with `drug_spec()` to see this in action:

```{r brand-example}
example_spec <- drug_spec(
  drug_class    = "example",
  label         = "Example GLP-1s",
  generic_names_probable = c("SEMAGLUTIDE", "TIRZEPATIDE"),
  generic_names_cautious = c("DULAGLUTIDE"),
  brand_names = list(
    SEMAGLUTIDE = "Ozempic",
    TIRZEPATIDE = "Mounjaro"
    # DULAGLUTIDE has no brand_names entry, so its brand is character(0)
  )
)

example_spec$get_generics(priority = 1:3)
```

Printing the tibble shows `<chr [1]>`/`<chr [0]>` for the list-column rather
than the brand text itself — that's expected; `str()` or indexing into the
column (e.g., `result$brand[[1]]`) shows the actual value(s). To get a
one-row-per-brand tibble instead, use `tidyr::unnest()` with
`keep_empty = TRUE` — **without** `keep_empty = TRUE`, `unnest()` silently
drops every row with no recorded brand, which is most of them:

```{r brand-unnest}
example_spec$get_generics(priority = 1:3) |>
  tidyr::unnest(brand, keep_empty = TRUE)
```

## Antidiabetic drug classes

`spec_antidiabetic` groups all oral antidiabetic agents and insulin:

```{r antidiab-spec}
spec_antidiabetic
```

```{r glp1}
get_antidiabetic_generics(component = "glp1_v1")
```

```{r sglt2}
get_antidiabetic_generics(component = "sglt2_v1")
```

## Creating your own drug spec

Use `drug_spec()` to define a custom drug class, or `modify_drug_spec()` to
extend an existing component. See `vignette("custom_specs")` for details.
