---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

```{r load}
library(peridefs)
```

`peridefs` provides code sets (ICD-9, ICD-10, and HCPCS/CPT) to define
conditions such as hypertension, diabetes, and atherosclerotic cardiovascular
disease in real world claims data. It also contains generic names for common
drug types (e.g., anti-diabetes). The purpose of the `peridefs` package is to
provide safe and easy access to these validated definitions.

## Retrieving condition codes

Each condition has one or more `get_<condition>_vX_codes()` functions — one
per algorithm version. By default the function returns a named list with one
element per code type:

```{r htn-list}
get_htn_v1_codes()
```

Pass `format = "tibble"` to get a tidy data frame instead

```{r htn-tibble}
get_htn_v1_codes(format = "tibble")
```

### Selecting a code type

By default, `get_` functions will return all relevant codes or generics. If you
only want to return a specific type or types of codes, use the `code_type`
argument to retrieve a filtered code list. Supprted values are `"dx_icd9"`,
`"dx_icd10"`, `"hcpcs"`, `"proc_icd9"`, and `"proc_icd10"`:

```{r htn-icd10}
get_htn_v1_codes(code_type = "dx_icd10")
```

Note that if you ask for a code type that isn't stored in the object, you get an empty list.

```{r}
get_htn_v1_codes(code_type = 'proc_icd10')
```


### Adding periods to codes

Codes are stored in short format (no periods, i.e., `.`) as this is a common
format in claims data. Set `periods = TRUE` to add them to your codes:

```{r htn-periods}
get_htn_v1_codes(code_type = "dx_icd10", periods = TRUE)
```

Note that this feature is not perfect - it just adds a period after the third character. Periods appear after the third character in ICD-10 but may appear after the fourth in ICD-9. Most of our analyses focus on ICD-10 and most databases do not use periods in codes, so getting this exactly right is not a high priority at the moment.

### Condition vs. outcome definitions

Some `specs` carry separate code sets for identifying a **condition** (history /
comorbidity lookback) vs. an **outcome** (new event). Use `variable_type` to
select one:

```{r hf-outcome}
get_hf_v1_codes(variable_type = "outcome", code_type = "dx_icd10")
```

The default (`variable_type = "condition"`) retrieves the condition definition.

## Composite specs and the `component` argument

Some specs are **composites** — they contain several named component specs
rather than a single code set. `spec_ascvd`, for example, groups coronary heart disease (CHD), stroke, lower extremity artery disease (LEAD) / peripheral arterial disease (PAD), and cerebrovascular disease into a single composition.

For composite specs, the `component` argument is **required**. Print the spec
to see all available component names:

```{r ascvd-spec}
spec_ascvd
```

Retrieve codes for a single component:

```{r ascvd-chd}
get_ascvd_codes(component = "chd_v1", code_type = "dx_icd10")
```

Use `component = "all"` to get the union of every component's codes:

```{r ascvd-all}
get_ascvd_codes(component = "all", code_type = "dx_icd10") |> length()
```

## Reading algorithm definitions

The codes are only half of what you need to get definitions right. You also need
to know how to apply them. `get_<condition>_vX_defs()` provides a description of
the algorithm used to define each condition:

```{r htn-defs}
get_htn_v1_defs()
```

For composite specs, pass the component name:

```{r ascvd-defs}
get_ascvd_defs(component = "chd_v1", variable_type = "condition")
```

## Versioning

Algorithm versions are encoded directly in the spec and function names using a
`_vX` suffix. `spec_htn_v1` and `spec_htn_v2` are two distinct objects; the
corresponding getters are `get_htn_v1_codes()` and `get_htn_v2_codes()`.
When a condition has only one version (e.g., `spec_hf_v1`), there is only one
getter.

```{r version-compare}
# v1 and v2 share the same code sets for HTN — only the defs differ
identical(get_htn_v1_codes(), get_htn_v2_codes())

get_htn_v1_defs()
get_htn_v2_defs()
```

## Spec objects

Behind every `get_*` function is a `spec_*` data object. You can inspect it
directly to see codes, definitions, and (for composites) component names:

```{r spec}
spec_htn_v1
```

# Complement with `icd` package

The `icd` package is not required for `peridefs`, but it can be used to complement it. In particular, if you want to look up the explanations of our diagnostic codes, `icd` is helpful and easy to apply:

```{r}

library(icd) # for as.icd9 / as.icd10 / explain_code functions
library(tidyverse)

peri_codes <- get_copd_v1_codes(code_type = 'dx_icd9')[[1]] 

as.icd9(peri_codes) %>% 
  set_names(peri_codes) %>% 
  map_chr(~explain_code(.x)) %>% 
  enframe(name = 'code', value = 'explanation')
  
```

Note this only works for diagnostic codes and `explain_code` sometimes fails to give explanations for all codes if you pass a vector (that's why I used map in my code).

## Learn more

See the [Reference](../reference/index.html) for the complete list of
conditions and drugs.
